---
title: "Bike Sharing Google Analytics CaseStudy"
output:
  word_document: default
  html_notebook: default
  pdf_document: default
---


#Case Dtudy : Bike Sharing

##Loading Packages:-


```{r}
library("tidyverse")    #Helps wrangle data
```

```{r}
library("lubridate")   #helps wrangle date attributes
```

```{r}
library("skimr")       #get summary data
library("janitor")
```

```{r}
library("dplyr")
```


#Step 1 Collect Data

```{r}
Trips_Apr20 <- read_csv('202004-divvy-tripdata.csv')
```

```{r}
Trips_May20 <- read_csv('202005-divvy-tripdata.csv')
```

```{r}
Trips_June20 <- read_csv('202006-divvy-tripdata.csv')
```

```{r}
Trips_July20 <- read_csv('202007-divvy-tripdata.csv')
```

```{r}
Trips_Aug20 <- read_csv('202008-divvy-tripdata.csv')
```

```{r}
Trips_Sep20 <- read_csv('202009-divvy-tripdata.csv')
```

```{r}
Trips_Oct20 <- read_csv('202010-divvy-tripdata.csv')
```

```{r}
Trips_Nov20 <- read_csv('202011-divvy-tripdata.csv')
```

```{r}
Trips_Dec20 <- read_csv('202012-divvy-tripdata.csv')
```

```{r}
Trips_Jan21 <- read_csv('202101-divvy-tripdata.csv')
```

```{r}
Trips_Feb21 <- read_csv('202102-divvy-tripdata.csv')
```

```{r}
Trips_March21 <- read_csv('202103-divvy-tripdata.csv')
```

```{r}
Trips_Apr21 <- read_csv('202104-divvy-tripdata.csv')
```

#Step 2 : Wrangle Data And Combine Into A Single File

###Compare columns names of the files. While the names don't have to be in the same order, they DO need to match perfectly ###before we can use a command to join them into one file.

```{r}
colnames(Trips_Apr20)
```



```{r}
colnames(Trips_May20)
```

```{r}
colnames(Trips_June20)
```

```{r}
colnames(Trips_July20)
```

```{r}
colnames(Trips_Aug20)
```

```{r}
colnames(Trips_Sep20)
```

```{r}
colnames(Trips_Oct20)
```

```{r}
colnames(Trips_Nov20)
```

```{r}
colnames(Trips_Dec20)
```

```{r}
colnames(Trips_Jan21)
```

```{r}
colnames(Trips_Feb21)
```

```{r}
colnames(Trips_March21)
```

```{r}
colnames(Trips_Apr21)
```

#Inspect the dataframes and look for inconguencies

```{r}
str(Trips_Apr20)
```

```{r}
str(Trips_May20)
```

```{r}
str(Trips_June20)
```

```{r}
str(Trips_July20)
```

```{r}
str(Trips_Aug20)
```
```{r}
str(Trips_Sep20)
```

```{r}
str(Trips_Oct20)
```

```{r}
str(Trips_Nov20)
```

```{r}
str(Trips_Dec20)
```

```{r}
str(Trips_Jan21)
```

```{r}
str(Trips_Feb21)
```

```{r}
str(Trips_March21)
```

```{r}
str(Trips_Apr21)
```

##We can compare column datatype across all dataframe by using compare_df_cols when we have large dataset, that would be more ##easy 

```{r}
compare_df_cols(Trips_Apr20,Trips_Apr21,Trips_Aug20,Trips_Dec20,Trips_Feb21,Trips_Jan21,Trips_July20,Trips_June20,Trips_March21,Trips_May20,Trips_Nov20,Trips_Sep20,Trips_Oct20, return = "mismatch")
```

##Convert end_station_id and start_station_id to character so that they can stack correctly

```{r}
Trips_Apr20 <- mutate(Trips_Apr20, end_station_id= as.character(end_station_id),start_station_id = as.character(start_station_id))


Trips_May20 <- mutate(Trips_May20, end_station_id= as.character(end_station_id),start_station_id = as.character(start_station_id))

Trips_June20 <- mutate(Trips_June20, end_station_id= as.character(end_station_id),start_station_id = as.character(start_station_id))

Trips_July20 <- mutate(Trips_July20, end_station_id= as.character(end_station_id),start_station_id = as.character(start_station_id))

Trips_Aug20 <- mutate(Trips_Aug20, end_station_id= as.character(end_station_id),start_station_id = as.character(start_station_id))

Trips_Sep20 <- mutate(Trips_Sep20, end_station_id= as.character(end_station_id),start_station_id = as.character(start_station_id))

Trips_Oct20 <- mutate(Trips_Oct20, end_station_id= as.character(end_station_id),start_station_id = as.character(start_station_id))

Trips_Nov20 <- mutate(Trips_Nov20, end_station_id= as.character(end_station_id),start_station_id = as.character(start_station_id))

Trips_Dec20 <- mutate(Trips_Dec20, end_station_id= as.character(end_station_id),start_station_id = as.character(start_station_id))

Trips_Jan21 <- mutate(Trips_Jan21, end_station_id= as.character(end_station_id),start_station_id = as.character(start_station_id))

Trips_Feb21 <- mutate(Trips_Feb21, end_station_id= as.character(end_station_id),start_station_id = as.character(start_station_id))

Trips_March21 <- mutate(Trips_March21, end_station_id= as.character(end_station_id),start_station_id = as.character(start_station_id))

Trips_Apr21 <- mutate(Trips_Apr21, end_station_id= as.character(end_station_id),start_station_id = as.character(start_station_id))

```

##Double Check column datatype across all dataframe

```{r}
compare_df_cols(Trips_Apr20,Trips_Apr21,Trips_Aug20,Trips_Dec20,Trips_Feb21,Trips_Jan21,Trips_July20,Trips_June20,Trips_March21,Trips_May20,Trips_Nov20,Trips_Sep20,Trips_Oct20, return = "mismatch")
```

##Stack individual data frames into one big data frame
```{r}
all_trips <- bind_rows(Trips_Apr20,Trips_Apr21,Trips_Aug20,Trips_Dec20,Trips_Feb21,Trips_Jan21,Trips_July20,Trips_June20,Trips_March21,Trips_May20,Trips_Nov20,Trips_Sep20,Trips_Oct20)
```

##Remove unused columns
```{r}
all_trips <- all_trips %>%
  select(-c(start_lat,start_lng,end_lat,end_lng))
```


##Rename Columns
```{r}
all_trips <- all_trips %>% 
  rename(trip_id = ride_id
         ,ride_type = rideable_type
         ,start_time = started_at
         ,end_time = ended_at
         ,from_station_name = start_station_name
         ,from_station_id = start_station_id
         ,to_station_name = end_station_name
         ,to_station_id = end_station_id
         ,usertype = member_casual
         )
```


#Step 3 : Clean Up and Add Data to prepare for analysis
##Inspect the new table that has been created

```{r}
colnames(all_trips)        #list of all columns of combined data
```

```{r}
dim(all_trips)            #Dianmensions of DataFrame
```

```{r}
head(all_trips)          #To See First 6 rows of dataframe
```

```{r}
str(all_trips)             #To see list of columns and datatypes(numeric, character, etc)
```

```{r}
summary(all_trips)             #Statistical summary of data, Manily for numerics
```

```{r}
skim(all_trips)           #To get summary of data, check missing data
```

#Now we will add columns that list the date, month, day, and year of each ride
##This will allow us to aggregate ride data for each month, day, or year....Before compleating these operations we could only aggregate at ride level

```{r}
all_trips$date <- as.Date(all_trips$start_time)     #The default format is yyyy-mm-dd

all_trips$month <- format(all_trips$date,"%m")

all_trips$day <- format(all_trips$date,"%d")

all_trips$year <- format(all_trips$date,"%y")

all_trips$day_of_week <- format(as.Date(all_trips$date),"%A")
```

#Add a "ride_length" calculation to all trips(in seconds)
```{r}
all_trips$ride_length <- difftime(all_trips$end_time,all_trips$start_time)
```

##Convert "ride_length" form factor to numeric so we can run calculation on the data
```{r}
is.factor(all_trips$ride_length)
```

```{r}
all_trips$ride_length <-as.numeric(as.character(all_trips$ride_length))

is.numeric(all_trips$ride_length)
```

#Remove Bad Data
##The Dataframe includes a few hundred entries when bikes were taken out of docks and checked for quality by Divvy or ride_length was negative
```{r}
skim(all_trips$ride_length)
```

```{r}
all_trips_v2 <- all_trips[!(all_trips$ride_length<0),]

skim(all_trips_v2)
```

#Step 4 : Conduct Descriptive analysis
###Descriptive analysis on ride_length(all figures in seconds)

```{r}
summary(all_trips_v2$ride_length)
```

#Export to CSV for further analysis
```{r}
write.csv(all_trips_v2,"data.csv")
```














Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Ctrl+Alt+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Ctrl+Shift+K* to preview the HTML file).

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.